Data Imputation as a Query Optimization

### 2016-12-05 paper/relatedwork.tex
paper: added related work

\section{Introduction}

>>>>

\section{Related Work}
\subsection{Missing Values and Statistics}
Imputation of missing values is a widely studied field within the statistics and machine
learning communities. As highlighted in \cite{gelman2006data}, missing data
can appear for a variety of reasons, including both random and conditioned on
existing values (observed and missing). Methods in the statistical community
focus on correctly modeling relationships between the attributes to factor in
such varied forms of missingness.

Burgette and Reiter introduce imputation of missing data using sequences
of regression trees in \cite{burgette2010multiple}. \ProjectName{} uses
this technique to impute missing values as needed in the generated query plan.
The flexibility of this non-parametric imputation approach aligns well with the
spirit of query optimization, which aims to reconcile performance with a declarative interface to
data manipulation. By removing concerns for imputation, users executing queries
in \ProjectName{} don't need to consider the potential nature of the missing data.

In \cite{akande2015empirical}, Akande et al analyze the performance of various
imputation techniques on American Community Survey data, "an ongoing survey
that provides vital information on a yearly basis"\cite{censusbureau}. The authors
use a common technique: multiple imputation, which imputes missing values
with various strategies on the base data, computes aggregates, and then
combines estimates of these aggregates to account for additional uncertainty across
strategies. The computational difficulties of imputing on large base tables
are well known and can limit approaches. For example, Akande finds that
one of their approaches (MI-GLM) crashes when attempting to impute on data that
includes variables with potentially large domains (ten categories specifically).
 In contrast, \ProjectName{} allows users to specify a tradeoff
between information loss and computational complexity (in terms of time). 
Furthermore, the query planner's imputation is guided by the requirements of
each specific query's operators, rather than requiring broad assumptions
about query workloads.

\subsection{Missing Values and Databases}
There is a long history in the database community surrounding the
treatment of nulls. As early as 1973, \cite{codd1973understanding}
provides a treatment of the semantics of null. Multiple
papers have described various (at times conflicting) treatments
of nulls \cite{grant1977null}. \ProjectName's main design invariant - that no relational operator
see attributes with missing values for attributes it must operate on - eliminates
the need to handle null value semantics, while guaranteeing soundness (modulo
imputation strategy) of the query evaluation.

BayesDB \cite{mansinghka2015bayesdb} provides users with a simple interface to 
leverage statistical inference techniques on a database. Non-experts
can use a simple declarative language, akin to SQL, to specify models
which allow missing value imputation, amongst other broader functionality.
Experts can further customize strategies and express domain knowledge to
improve performance and accuracy.

While BayesDB can be used for value imputation, this step is not framed
within the context of query planning, but rather as an explicit step, using the
\verb|INFER| operation.

BayesDB provides a great alternative for bridging the gap between
traditional databases and sophisticated modeling software. \ProjectName{}, in
contrast, aims to remain squarely in the database realm, while allowing
users to perform queries on a potentially larger subset of their data.

\ProjectName's cost-based query planner 
is partially based on the seminal work developed for System R's query planning\cite{blasgen1981system}.
However, in contrast to System R, \ProjectName{} performs additional histogram transformations to account
for the changing nature of missing values.
